---
title: "p8105_hw6_yj2686"
author: "Yiqun Jin"
date: "11/30/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
set.seed(1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Problem 1

### Load and clean the data for regression analysis

```{r birthweight}
birthweight = read.csv("./birthweight.csv") %>% 
  mutate(babysex = factor(if_else(babysex == 1, "male", "female")),
         frace = factor(recode(frace, '1' = "White", '2' = "Black", '3' = "Asian", 
                               '4' = "Puerto Rican", '8' = "Other", '9' = "Unknown")),
         mrace = factor(recode(mrace,'1' = "White", '2' = "Black", '3' = "Asian", 
                               '4' = "Puerto Rican", '8' = "Other", '9' = "Unknown")),
         malform = factor(recode(malform, '0' = "absent", '1' = "present")))
# check for missing data
sum(is.na(birthweight))
```

### Propose a regression model for birth weight.

Since birth weight of baby is usually associated with baby's length at birth, I made a scatter plot of baby's length at birth `blength`. 

```{r}
birthweight %>% 
  ggplot(aes(x = blength, y = bwt)) + 
  geom_point(alpha = .5) +
  labs(title = "Scatter plot of baby's birth weight and baby's birth length") +
  theme(plot.title = element_text(hjust = 0.5))
```

It seemed that there might be a linear relationship between baby's length at birth `blength` and birth weight `bwt`. Thus, I will try `blength` as the predictor of the regression model for birth weight and make a plot of model residuals against fitted values

```{r}
fit1 = lm(bwt ~ blength, data = birthweight)
fit1 %>% 
  broom::tidy() %>% 
  knitr::kable()

birthweight %>% 
  modelr::add_residuals(fit1) %>% 
  modelr::add_predictions(fit1) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(title = "Plot of model residuals against fitted values for fit1 model",
       x = "Fitted Values",
       y = "Residuals") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
fit2 = lm(bwt ~ blength + gaweeks,birthweight)

fit2 %>% 
  broom::tidy() %>% 
  knitr::kable()

birthweight %>% 
  modelr::add_residuals(fit2) %>% 
  modelr::add_predictions(fit2) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(title = "Plot of model residuals against fitted values for fit2 model",
       x = "Fitted Values",
       y = "Residuals") +
  theme(plot.title = element_text(hjust = 0.5))

fit3 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, birthweight)

fit3 %>% 
  broom::tidy() %>% 
  knitr::kable()

birthweight %>% 
  modelr::add_residuals(fit3) %>% 
  modelr::add_predictions(fit3) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(title = "Plot of model residuals against fitted values for fit3 model",
       x = "Fitted Values",
       y = "Residuals") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Make this comparison in terms of the cross-validated prediction error

```{r}
cv_df = 
  crossv_mc(birthweight, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  ) %>% 
  mutate(
    fit1_mod = map(train, ~lm(bwt ~ blength, data = birthweight)),
    fit2_mod = map(train, ~lm(bwt ~ blength + gaweeks,birthweight)),
    fit3_mod = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex +       blength*babysex + bhead*blength*babysex, birthweight))
  ) %>% 
  mutate(
    rmse_fit1 = map2_dbl(fit1_mod, test, ~rmse(model = .x, data = .y)),
    rmse_fit2 = map2_dbl(fit2_mod, test, ~rmse(model = .x, data = .y)),
    rmse_fit3 = map2_dbl(fit3_mod, test, ~rmse(model = .x, data = .y))
  )
  
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  labs(
    title = "distribution of RMSE values for three candidate model",
    y = "Root-mean-square-deviation"
  )+
  theme(plot.title = element_text(hjust = 0.5))
```

According to the distribution of RMSE, fit3 model have the lowest RMSE value, meaning that fit3 would be the most optimal model among three candidate models


## Problem 2

```{r weather data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Bootstrapping

```{r}
bootstrap = 
  weather_df %>% 
  modelr::bootstrap(n = 5000, id = "strap_id") %>% 
  mutate(
    strap = map(strap, as_tibble),
    models = map(strap, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy),
    summary = map(models, broom::glance)
  ) %>% 
  unnest(c(results, summary), names_repair = "universal")
```

