---
title: "p8105_hw6_yj2686"
author: "Yiqun Jin"
date: "11/30/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
set.seed(1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Problem 1

Load and clean the data for regression analysis

```{r birthweight}
birthweight = read.csv("./birthweight.csv") %>% 
  mutate(babysex = factor(if_else(babysex == 1, "male", "female")),
         frace = factor(recode(frace, '1' = "White", '2' = "Black", '3' = "Asian", 
                               '4' = "Puerto Rican", '8' = "Other", '9' = "Unknown")),
         mrace = factor(recode(mrace,'1' = "White", '2' = "Black", '3' = "Asian", 
                               '4' = "Puerto Rican", '8' = "Other", '9' = "Unknown")),
         malform = factor(recode(malform, '0' = "absent", '1' = "present")))
# check for missing data
sum(is.na(birthweight))
```

Propose a regression model for birth weight.

Since birth weight of baby is usually associated with baby's length at birth, I made a scatter plot of baby's length at birth `blength`. 

```{r}
birthweight %>% 
  ggplot(aes(x = blength, y = bwt)) + 
  geom_point(alpha = .5) +
  labs(title = "Scatter plot of baby's birth weight and baby's length at birth") +
  theme(plot.title = element_text(hjust = 0.5))
```

It seemed that there might be a linear relationship between baby's length at birth `blength` and birth weight `bwt`. Thus, I will try `blength` as the predictor of the regression model for birth weight and make a plot of model residuals against fitted values

```{r}
fit1 = lm(bwt ~ blength, data = birthweight)
fit1 %>% 
  broom::tidy() %>% 
  knitr::kable()

birthweight %>% 
  modelr::add_residuals(fit1) %>% 
  modelr::add_predictions(fit1) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .3) +
  labs(title = "Plot of model residuals against fitted values",
       x = "Fitted Values",
       y = "Residuals") +
  theme(plot.title = element_text(hjust = 0.5))
```

Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

```{r}
fit2 = lm(bwt ~ blength + gaweeks,birthweight)
fit3 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, birthweight)

fit2 %>% 
  broom::tidy() %>% 
  knitr::kable()

fit3 %>% 
  broom::tidy() %>% 
  knitr::kable()
```

## Problem 2

```{r weather data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Bootstrapping

```{r}
bootstrap = 
  weather_df %>% 
  modelr::bootstrap(n = 5000, id = "strap_id") %>% 
  mutate(
    strap = map(strap, as_tibble),
    models = map(strap, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy),
    summary = map(models, broom::glance)
  ) %>% 
  unnest(c(results, summary), names_repair = "universal")
```

